%************************************************
% ACKNOWLEDGEMENTS
%************************************************

\begin{acknowledgements}
I would like to sincerely thank my supervisor Prof. Cesare Alippi for giving me the opportunity to work on such an interesting project and my co-supervisor Dr. Daniele Grattarola for giving me all the support I needed during this project and for allowing me to learn a lot from him. I couldn't have asked for better guidance.

A huge thanks goes to my parents, which gave me the chance to attend this great university in Lugano and to achieve my goals by supporting and encouraging me like no one else could. Thanks also to my brother Andrea and my sister Roberta for their unique way to reassure me and make me feel their support. Thanks to all my relatives, which made me feel their presence even from far away and, in particular, to my aunt Alessandra and my cousin Valentina for always taking to heart my situations and helping me whenever I needed.

I want to say a big thank you to all my friends from Bologna, Milano and Lugano. Thanks to those who were close for the technical and moral support during the last two years and to those who were distant for making me feel at home every time I came back, as if no time had passed. Thanks to Francesco for making me laugh even during difficult moments at university and for teaching me a lot of stuff in the field we are both passionate about. Thanks to Francesca Z., Valentina O., Francesca T., Giulia T. and Alessandro R. for always being an essential part of my life even from far away and for advising and supporting me in the best way I could ever hope for. Thanks to my bachelor buddies in Milano for always welcoming me as part of a family.

Finally, thank you to Roberto for representing my reference point during the last two years and for encouraging me in the most effective ways. Thank you for being by my side and for your patience. Thank you for always being present when I needed the most and for celebrating with me my achievements. Thank you for sharing your life with me and for always having me in your thoughts. I feel lucky and grateful for all you do for me and you'll always be my main source of \textit{nefer}.
\end{acknowledgements}